Modelling state-switching functional data with hidden Markov models: supplement A (case study)

### 2021-06-21 Report/Supp_Case.tex
Updates from Overleaf

% Title, authors, affiliations
\title{Modelling state-switching functional data with hidden Markov models: supplementary material (case study)}%\query{Q1}
\date{}
\author{BlindedA, BlindedB}%\authorref{1}\thanksref{*}}
%\author{BlindedB}%\authorref{2}}
%\affiliation[1]{Author affiliations will go here in the accepted manuscript, but do NOT include them in your initial submission because it must be anonymous.}
%\affiliation[2]{Second Affiliation}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the dive duration parameters $\mu_Y^{(i)}$ and $\sigma_Y^{(i)}$ for dive types $i = 1,2$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is scaled by the log-likelihood of the maximum likelihood estimate $(\hat \mu_Y,\hat \sigma_Y)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the coarse-scale probability transition matrix parameters $\eta_{1,2}$ and $\eta_{2,1}$. Recall that $\Gamma_{i,j} = \exp(\eta_{i,j}) / \sum_k \exp(\eta_{i,k})$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is scaled by the log-likelihood of the maximum likelihood estimate $(\hat \eta_{1,2},\hat \eta_{2,1})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the wiggliness parameters $\mu_W^{*(\cdot,i^*)}$ and $\sigma_W^{*(\cdot,i^*)}$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is scaled by the log-likelihood of the maximum likelihood estimate $(\hat \mu_W^{*(\cdot,i^*)}, \hat \sigma_W^{*(\cdot,i^*)})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $x$-acceleration parameters $\mu_{A_x}^{*(\cdot,i^*)}, \sigma_{A_x}^{*(\cdot,i^*)}$ and $\text{logit}\left(\phi_{A}^{*(\cdot,i^*)}\right)$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is scaled by the log-likelihood of the maximum likelihood estimate $\left(\hat \mu_{A_x}^{*(\cdot,i^*)}, \hat \sigma_{A_x}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)}\right)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $y$-acceleration parameters $\mu_{A_y}^{*(\cdot,i^*)}, \sigma_{A_y}^{*(\cdot,i^*)}$ and $\text{logit}\left(\phi_{A}^{*(\cdot,i^*)}\right)$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is scaled by the log-likelihood of the maximum likelihood estimate $\left(\hat \mu_{A_y}^{*(\cdot,i^*)}, \hat \sigma_{A_y}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)}\right)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $z$-acceleration parameters $\mu_{A_z}^{*(\cdot,i^*)}, \sigma_{A_z}^{*(\cdot,i^*)}$ and $\text{logit}\left(\phi_{A}^{*(\cdot,i^*)}\right)$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is scaled by the log-likelihood of the maximum likelihood estimate $\left(\hat \mu_{A_z}^{*(\cdot,i^*)}, \hat \sigma_{A_z}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)}\right)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the fine-scale probability transition matrix parameters associated both dive types ($\eta^{(i)*}_{i^*,j^*}$ for $i = 1,2$ and $i^*,j^* = 1,2,3$). Recall that $\eta^{(i)*}_{i^*,i^*} = 0$ and that $\Gamma^{(i)*}_{i^*,j^*} = \exp(\eta^{(i)*}_{i^*,j^*}) / \sum_{k^*} \exp(\eta^{(i)*}_{i^*,k^*})$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is scaled by the log-likelihood of the maximum likelihood estimate $\hat \Gamma^{(i)*}$, which is denoted by a red star. Note that the long ridges of approximately equal likelihood corresponds to $\hat \Gamma^{(i)*}_{1,3} \approx \hat \Gamma^{(i)*}_{3,1} \approx 0$. This indicates that the whale never switches directly between subdive states 1 and 3.
        \addtocounter{fignum}{1}

@@

            %\begin{table}[ht]
            %
            %\caption{Estimates and empirical standard errors of emission parameters for dive duration ($Y_t$), acceleration ($\Zone_{t,\tilde t^*}$), and wiggliness ($\Ztwo_{t,\tilde t^*}$) of the killer whale case study data under the full CarHHMM-DFT.}
            \begin{center}
            \scalebox{0.75}{
            \bgroup
            \centering
            \def\arraystretch{1.5}
            \begin{tabular}{ccccc}
                \multirow{2}{*}{Feature}                                                       & \multirow{2}{*}{Dive / Subdive Type} & \multicolumn{3}{c}{Parameter Estimate}        \\
                                                                                               &                                      & $\hat \mu$    & $\hat \sigma$ & $\hat \phi$   \\ \hline
                \multirow{2}{*}{Dive Duration $(s)$ -- $Y_t$}                                  & 1                                    & $27.34 (0.63)$ & $10.96 (0.56)$ & ---           \\
                                                                                               & 2                                    & $127.5 (11.3)$ & $63.9 (9.0)$ & ---           \\ \hline
                \multirow{3}{*}{$x$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_x$}   & 1                                    & $0.449 (0.030)$ & $0.039 (0.001)$ & $0.968 (0.002)$ \\
                                                                                               & 2                                    & $0.210 (0.012)$ & $0.096 (0.002)$ & $0.829 (0.010)$ \\
                                                                                               & 3                                    & $0.232 (0.035)$ & $0.296 (0.010)$ & $0.607 (0.042)$ \\ \hline
                \multirow{3}{*}{$y$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_y$}   & 1                                    & $0.450 (0.038)$ & $0.051 (0.001)$ & $0.968 (0.009)$ \\
                                                                                               & 2                                    & $0.437 (0.012)$ & $0.094 (0.002)$ & $0.829 (0.016)$ \\
                                                                                               & 3                                    & $0.366 (0.042)$ & $0.365 (0.012)$ & $0.607 (0.040)$ \\ \hline
                \multirow{3}{*}{$z$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_z$}   & 1                                    & $-0.691 (0.043)$ & $0.058 (0.001)$ & $0.968 (0.005)$ \\
                                                                                               & 2                                    & $-0.573 (0.014)$ & $0.111 (0.002)$ & $0.829 (0.012)$ \\
                                                                                               & 3                                    & $-0.303 (0.041)$ & $0.354 (0.012)$ & $0.607 (0.039)$ \\ \hline
                \multirow{3}{*}{Wiggliness - $\Ztwo_{t,\tilde t^*}$}                           & 1                                    & $34.02 (0.36)$ & $22.99 (0.38)$ & ---           \\
                                                                                               & 2                                    & $490.1 (5.6)$ & $502.6 (6.8)$ & ---           \\
                                                                                               & 3                                    & $9150 (220)$ & $13540 (350)$ & ---           \\ \hline
            \end{tabular}
            \egroup
            }
            \end{center}

@@

        \subsection{HHMM-DFT}
            \begin{center}
            \scalebox{0.75}{
                \bgroup
                \centering
                \def\arraystretch{1.5}
                \begin{tabular}{ccccc}
                \multirow{2}{*}{Feature}                                                       & \multirow{2}{*}{Dive / Subdive Type} & \multicolumn{3}{c}{Parameter Estimate}          \\
                                                                                               &                                      & $\hat \mu$    & $\hat \sigma$ & $\hat \phi$     \\ \hline
                \multirow{2}{*}{Dive Duration $(s)$ -- $Y_t$}                                  & 1                                    & $26.61 (0.61)$ & $10.16 (0.52)$ & ---             \\
                                                                                               & 2                                    & $116.8 (10.6)$ & $64.7 (8.6)$ & ---             \\ \hline
                \multirow{3}{*}{$x$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_x$}   & 1                                    & $0.000 (0.002)$ & $0.065 (0.001)$ & ---             \\
                                                                                               & 2                                    & $0.085 (0.004)$ & $0.175 (0.003)$ & ---             \\
                                                                                               & 3                                    & $0.291 (0.020)$ & $0.484 (0.014)$ & ---             \\ \hline
                \multirow{3}{*}{$y$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_y$}   & 1                                    & $0.365 (0.002)$ & $0.068 (0.001)$ & ---             \\
                                                                                               & 2                                    & $0.408 (0.002)$ & $0.108 (0.002)$ & ---             \\
                                                                                               & 3                                    & $0.275 (0.017)$ & $0.407 (0.012)$ & ---             \\ \hline
                \multirow{3}{*}{$z$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_z$}   & 1                                    & $-0.442 (0.003)$ & $0.101 (0.002)$ & ---             \\
                                                                                               & 2                                    & $-0.471 (0.003)$ & $0.158 (0.002)$ & ---             \\
                                                                                               & 3                                    & $-0.298 (0.017)$ & $0.415 (0.012)$ & ---             \\ \hline
                \multirow{3}{*}{Wiggliness - $\Ztwo_{t,\tilde t^*}$}                           & 1                                    & $43.65 (0.52)$ & $33.66 (0.58)$ & ---             \\
                                                                                               & 2                                    & $608.0 (6.8)$ & $717.3 (9.2)$ & ---             \\
                                                                                               & 3                                    & $7100 (160)$ & $13620 (310)$ & ---             \\ \hline
                \end{tabular}
                \egroup
            }
            \end{center}

@@

        \subsection{CarHMM-DFT}
            \begin{center}
            \scalebox{0.75}{
                \bgroup
                \centering
                \def\arraystretch{1.5}
            \begin{tabular}{ccccc}
            \multirow{2}{*}{Feature}                                                       & \multirow{2}{*}{Dive / Subdive Type} & \multicolumn{3}{c}{Parameter Estimate}        \\
                                                                                           &                                      & $\hat \mu$    & $\hat \sigma$ & $\hat \phi$   \\ \hline
            \multirow{1}{*}{Dive Duration $(s)$ -- $Y_t$}                                  & 1                                    & $41.70 (1.24)$ & $30.50 (1.22)$ & ---           \\ \hline
            \multirow{3}{*}{$x$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_x$}   & 1                                    & $0.460 (0.030)$ & $0.039 (0.001)$ & $0.968 (0.002)$ \\
                                                                                           & 2                                    & $0.209 (0.012)$ & $0.096 (0.002)$ & $0.828 (0.007)$ \\
                                                                                           & 3                                    & $0.232 (0.035)$ & $0.296 (0.010)$ & $0.607 (0.023)$ \\ \hline
            \multirow{3}{*}{$y$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_y$}   & 1                                    & $0.457 (0.038)$ & $0.052 (0.001)$ & $0.968 (0.002)$ \\
                                                                                           & 2                                    & $0.436 (0.012)$ & $0.094 (0.002)$ & $0.828 (0.007)$ \\
                                                                                           & 3                                    & $0.365 (0.042)$ & $0.366 (0.012)$ & $0.607 (0.023)$ \\ \hline
            \multirow{3}{*}{$z$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_z$}   & 1                                    & $-0.704 (0.044)$ & $0.059 (0.001)$ & $0.968 (0.002)$ \\
                                                                                           & 2                                    & $-0.571 (0.014)$ & $0.111 (0.002)$ & $0.828 (0.007)$ \\
                                                                                           & 3                                    & $-0.301 (0.042)$ & $0.355 (0.012)$ & $0.607 (0.023)$ \\ \hline
            \multirow{3}{*}{Wiggliness - $\Ztwo_{t,\tilde t^*}$}                           & 1                                    & $34.40 (0.37)$ & $23.33 (0.38)$ & ---           \\
                                                                                           & 2                                    & $495.0 (5.7)$ & $505.4 (6.9)$ & ---           \\
                                                                                           & 3                                    & $9190 (220)$ & $13590 (360)$ & ---           \\ \hline
            \end{tabular}
                \egroup
            }
            \end{center}

                \noindent Table \arabic{tablenum}: Estimates and predicted standard errors of emission parameters for dive duration ($Y_t$), acceleration ($\Zone_{t,\tilde t^*}$), and wiggliness ($\Ztwo_{t,\tilde t^*}$) of the killer whale case study data under the CarHMM-DFT. The parentheses refer to the estimated standard error of the parameter estimates derived from the inverse of the observed information matrix. Note that $\hat \mu_A^{*(\cdot,1)}$ is undefined because $\hat \phi_A^{*(\cdot,1)} = 1$, indicating that the mean of $\Zone_{t,\tilde t^*}$ is equal to $\Zone_{t,\tilde t^*-1}$ when in subdive state 1.
                \addtocounter{tablenum}{1}

@@

\iffalse
\newpage
%\section{Simulation Study Results}

    \section{Simulated Data}

        \begin{center}
    	\includegraphics[width=5in]{../Plots/sim_data.png}
    	\end{center}

    	\noindent Figure \arabic{fignum}: Acceleration data for a five dives of a selected simulated data set. The raw accelerometer data is denoted as $Y^*_{t,\tilde t^*}$ which is recorded at 50 hertz while $\Zone_{t,\tilde t^*}$ and $\Ztwo_{t,\tilde t^*}$ are recorded at 0.5 hertz. Dives are separated by vertical black lines. The colour of the curve corresponds to the true fine-scale state, while the colour of the background corresponds to the true dive type. Subdive state 1 can plausibly be interpreted as ``gliding" while subdive state 2 can plausibly be interpreted as active swimming, or ``fluking".
    	\addtocounter{fignum}{1}

    \newpage
    \section{Accuracy comparisons}

        \begin{center}
        \scalebox{0.75}{
        \bgroup
        \def\arraystretch{1.5}
        \begin{tabular}{ccccccc}
Model                       & \multicolumn{1}{c}{Train Time (m)} & \multicolumn{1}{c}{Dive Type} & \multicolumn{1}{c}{Subdive Type} & \multicolumn{1}{c}{Dive Accuracy} & \multicolumn{1}{c}{Subdive Accuracy}  \\ \hline
\multirow{5}{*}{CarHMM-DFT} & \multirow{5}{*}{$70 \pm 11$}   & Both                          & Both                             & -------------                     & $0.934 \pm 0.005$                       \\
                            &                                    & 1                             & 1                                & \multirow{2}{*}{-------------}    & $0.794 \pm 0.037$                       \\ 
                            &                                    & 1                             & 2                                &                                   & $0.958 \pm 0.008$                       \\ 
                            &                                    & 2                             & 1                                & \multirow{2}{*}{-------------}    & $0.924 \pm 0.012$                       \\ 
                            &                                    & 2                             & 2                                &                                   & $0.941 \pm 0.011$                       \\ \hline 
\multirow{5}{*}{HHMM-DFT}   & \multirow{5}{*}{$209 \pm 51$}   & Both                          & Both                             & $0.942 \pm 0.036$                   & $0.881 \pm 0.041$                       \\
                            &                                    & 1                             & 1                                & \multirow{2}{*}{$0.966\pm0.030$}    & $0.628 \pm 0.120$                       \\ 
                            &                                    & 1                             & 2                                &                                   & $0.956 \pm 0.014$                       \\ 
                            &                                    & 2                             & 1                                & \multirow{2}{*}{$0.852\pm0.095$}    & $0.789 \pm 0.158$                       \\ 
                            &                                    & 2                             & 2                                &                                   & $0.918 \pm 0.028$                       \\ \hline
\multirow{5}{*}{CarHHMM}    & \multirow{5}{*}{$236 \pm 52$}   & Both                          & Both                             & $0.873 \pm 0.165$                   & $0.740 \pm 0.051$                       \\
                            &                                    & 1                             & 1                                & \multirow{2}{*}{$0.886\pm0.214$}    & $0.765 \pm 0.205$                       \\ 
                            &                                    & 1                             & 2                                &                                   & $0.707 \pm 0.090$                       \\ 
                            &                                    & 2                             & 1                                & \multirow{2}{*}{$0.823\pm0.136$}    & $0.873 \pm 0.225$                       \\ 
                            &                                    & 2                             & 2                                &                                   & $0.657 \pm 0.099$                       \\ \hline
\multirow{5}{*}{CarHHMM-DFT}& \multirow{5}{*}{$132 \pm 40$}   & Both                          & Both                             & $0.944 \pm 0.036$                   & $0.934 \pm 0.005$                       \\
                            &                                    & 1                             & 1                                & \multirow{2}{*}{$0.963\pm0.031$}    & $0.759 \pm 0.042$                       \\ 
                            &                                    & 1                             & 2                                &                                   & $0.962 \pm 0.008$                       \\ 
                            &                                    & 2                             & 1                                & \multirow{2}{*}{$0.872\pm0.086$}    & $0.934 \pm 0.010$                       \\ 
                            &                                    & 2                             & 2                                &                                   & $0.932 \pm 0.011$                       \\ \hline
\end{tabular}
        \egroup
        }
        \end{center}

        \noindent Table \arabic{tablenum}: Average decoding accuracies and training times for all models used to categorize dive type and subdive state in the simulation study. Each of the four models was fit to 500 training data sets comprised of 100 simulated dives and tested on test data sets also comprised of 100 simulated dives. Reported values are averages, and $\pm$ refers to the sample standard deviation across the 500 data sets. Rows labelled Both/Both correspond to overall average decoding accuracy.
        \addtocounter{tablenum}{1}

    \newpage
    \section{Empirical joint distributions - dive duration ($Y_t$)}

        \begin{center}
        \scalebox{0.75}{
        \bgroup
        \centering
        \def\arraystretch{1.5}
\begin{tabular}{ccccccc}
Model                       & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Dive Type} & \multicolumn{1}{c}{Estimate}   & \multicolumn{1}{c}{Bias}   & \multicolumn{1}{c}{Empirical SE}   & \multicolumn{1}{c}{Observed Fischer SE}     \\ \hline
\multirow{4}{*}{CarHMM-DFT} & \multirow{2}{*}{$\mu$}        & 1                             & $42.161$                         & ---          & $4.255$                             & $2.036 \pm 0.369$                             \\
                            &                               & ---                           & ---                            & ---                        & ---                                & ---                                         \\
                            & \multirow{2}{*}{$\sigma$}     & 1                             & $31.119$                         & ---          & $5.393$                             & $2.014 \pm 0.452$                             \\
                            &                               & ---                           & ---                            & ---                        & ---                                & ---                                         \\ \hline
\multirow{4}{*}{HHMM-DFT}   & \multirow{2}{*}{$\mu$}        & 1                             & $25.776$                         & $0.096$ $(p=0.071)$          & $1.181$                             & $0.959 \pm 0.111$                             \\
                            &                               & 2                             & $108.589$                         & $3.989$ $(p=0.000)$          & $18.726$                             & $13.083 \pm 4.546$                             \\

>>>>

% Title, authors, affiliations
\title{Modelling state-switching functional data with hidden Markov models: supplement A  (case study)}%\query{Q1}
\date{}
\author{BlindedA, BlindedB}%\authorref{1}\thanksref{*}}
%\author{BlindedB}%\authorref{2}}
%\affiliation[1]{Author affiliations will go here in the accepted manuscript, but do NOT include them in your initial submission because it must be anonymous.}
%\affiliation[2]{Second Affiliation}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the dive duration parameters $\mu_Y^{(i)}$ and $\sigma_Y^{(i)}$ for dive types $i = 1,2$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \mu_Y,\hat \sigma_Y)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the coarse-scale probability transition matrix parameters $\eta_{1,2}$ and $\eta_{2,1}$. Recall that $\Gamma_{i,j} = \exp(\eta_{i,j}) / \sum_k \exp(\eta_{i,k})$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \eta_{1,2},\hat \eta_{2,1})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the wiggliness parameters $\mu_W^{*(\cdot,i^*)}$ and $\sigma_W^{*(\cdot,i^*)}$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $(\hat \mu_W^{*(\cdot,i^*)}, \hat \sigma_W^{*(\cdot,i^*)})$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $x$-acceleration parameters $\mu_{A_x}^{*(\cdot,i^*)}, \sigma_{A_x}^{*(\cdot,i^*)}$ and $\text{logit}\left(\phi_{A}^{*(\cdot,i^*)}\right)$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $\left(\hat \mu_{A_x}^{*(\cdot,i^*)}, \hat \sigma_{A_x}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)}\right)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $y$-acceleration parameters $\mu_{A_y}^{*(\cdot,i^*)}, \sigma_{A_y}^{*(\cdot,i^*)}$ and $\text{logit}\left(\phi_{A}^{*(\cdot,i^*)}\right)$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $\left(\hat \mu_{A_y}^{*(\cdot,i^*)}, \hat \sigma_{A_y}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)}\right)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the $z$-acceleration parameters $\mu_{A_z}^{*(\cdot,i^*)}, \sigma_{A_z}^{*(\cdot,i^*)}$ and $\text{logit}\left(\phi_{A}^{*(\cdot,i^*)}\right)$ for subdive types $i^* = 1,2,3$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $\left(\hat \mu_{A_z}^{*(\cdot,i^*)}, \hat \sigma_{A_z}^{*(\cdot,i^*)}, \hat \phi_{A}^{*(\cdot,i^*)}\right)$, which is denoted by a red star.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Log-likelihood of the CarHHMM-DFT as a function of the fine-scale probability transition matrix parameters associated both dive types ($\eta^{*(i)}_{i^*,j^*}$ for dive types $i = 1,2$ and subdive states $i^*,j^* = 1,2,3$). Recall that $\eta^{*(i)}_{i^*,i^*} = 0$ and that $\Gamma^{*(i)}_{i^*,j^*} = \exp(\eta^{*(i)}_{i^*,j^*}) / \sum_{k^*} \exp(\eta^{*(i)}_{i^*,k^*})$. All other parameters and probability transition matrices are assumed to be the maximum likelihood estimates shown in Table (1) of Section 3.1. The log-likelihood is added to the negative log-likelihood of the maximum likelihood estimate $\hat \Gamma^{*(i)}$, which is denoted by a red star. Note that the long ridges of approximately equal likelihood are due to the fact that $\hat \Gamma^{*(i)}_{1,3} \approx \hat \Gamma^{*(i)}_{3,1} \approx 0$. This indicates that the whale rarely switches directly between subdive states 1 and 3.
        \addtocounter{fignum}{1}

@@

            %\begin{table}[ht]
            %
            %\caption{Estimates and empirical standard errors of emission parameters for dive duration ($Y_t$), acceleration ($\Zone_{t,\tilde t^*}$), and wiggliness ($\Ztwo_{t,\tilde t^*}$) of the killer whale case study data under the full CarHHMM-DFT.}
            \begin{center}
            \scalebox{0.75}{
            \bgroup
            \centering
            \def\arraystretch{1.5}
            \begin{tabular}{ccccc}
                \multirow{2}{*}{Feature}                                                       & \multirow{2}{*}{Dive / Subdive Type} & \multicolumn{3}{c}{Parameter Estimate}        \\
                                                                                               &                                      & $\hat \mu$    & $\hat \sigma$ & $\hat \phi$   \\ \hline
                \multirow{2}{*}{Dive Duration $(s)$ -- $Y_t$}                                  & 1                                    & $27.34 (0.63)$ & $10.96 (0.56)$ & ---           \\
                                                                                               & 2                                    & $127.5 (11.3)$ & $63.9 (9.0)$ & ---           \\ \hline
                \multirow{3}{*}{$x$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_x$}   & 1                                    & $0.449 (0.030)$ & $0.039 (0.001)$ & $0.968 (0.002)$ \\
                                                                                               & 2                                    & $0.210 (0.012)$ & $0.096 (0.002)$ & $0.829 (0.007)$ \\
                                                                                               & 3                                    & $0.232 (0.035)$ & $0.296 (0.010)$ & $0.607 (0.023)$ \\ \hline
                \multirow{3}{*}{$y$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_y$}   & 1                                    & $0.450 (0.038)$ & $0.051 (0.001)$ & $0.968 (0.002)$ \\
                                                                                               & 2                                    & $0.437 (0.012)$ & $0.094 (0.002)$ & $0.829 (0.007)$ \\
                                                                                               & 3                                    & $0.366 (0.042)$ & $0.365 (0.012)$ & $0.607 (0.023)$ \\ \hline
                \multirow{3}{*}{$z$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_z$}   & 1                                    & $-0.691 (0.043)$ & $0.058 (0.001)$ & $0.968 (0.002)$ \\
                                                                                               & 2                                    & $-0.573 (0.014)$ & $0.111 (0.002)$ & $0.829 (0.007)$ \\
                                                                                               & 3                                    & $-0.303 (0.041)$ & $0.354 (0.012)$ & $0.607 (0.023)$ \\ \hline
                \multirow{3}{*}{Wiggliness - $\Ztwo_{t,\tilde t^*}$}                           & 1                                    & $34.02 (0.36)$ & $22.99 (0.38)$ & ---           \\
                                                                                               & 2                                    & $490.1 (5.6)$ & $502.6 (6.8)$ & ---           \\
                                                                                               & 3                                    & $9150 (220)$ & $13540 (350)$ & ---           \\ \hline
            \end{tabular}
            \egroup
            }
            \end{center}

@@

            \begin{center}
            \scalebox{0.75}{
                \bgroup
                \centering
                \def\arraystretch{1.5}
                \begin{tabular}{ccccc}
                \multirow{2}{*}{Feature}                                                       & \multirow{2}{*}{Dive / Subdive Type} & \multicolumn{3}{c}{Parameter Estimate}          \\
                                                                                               &                                      & $\hat \mu$    & $\hat \sigma$ & $\hat \phi$     \\ \hline
                \multirow{2}{*}{Dive Duration $(s)$ -- $Y_t$}                                  & 1                                    & $26.61 (0.61)$ & $10.16 (0.52)$ & ---             \\
                                                                                               & 2                                    & $116.8 (10.6)$ & $64.7 (8.6)$ & ---             \\ \hline
                \multirow{3}{*}{$x$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_x$}   & 1                                    & $0.000 (0.002)$ & $0.065 (0.001)$ & ---             \\
                                                                                               & 2                                    & $0.085 (0.004)$ & $0.175 (0.003)$ & ---             \\
                                                                                               & 3                                    & $0.291 (0.020)$ & $0.484 (0.014)$ & ---             \\ \hline
                \multirow{3}{*}{$y$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_y$}   & 1                                    & $0.365 (0.002)$ & $0.068 (0.001)$ & ---             \\
                                                                                               & 2                                    & $0.408 (0.002)$ & $0.108 (0.002)$ & ---             \\
                                                                                               & 3                                    & $0.275 (0.017)$ & $0.407 (0.012)$ & ---             \\ \hline
                \multirow{3}{*}{$z$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_z$}   & 1                                    & $-0.442 (0.003)$ & $0.101 (0.002)$ & ---             \\
                                                                                               & 2                                    & $-0.471 (0.003)$ & $0.158 (0.002)$ & ---             \\
                                                                                               & 3                                    & $-0.298 (0.017)$ & $0.415 (0.012)$ & ---             \\ \hline
                \multirow{3}{*}{Wiggliness - $\Ztwo_{t,\tilde t^*}$}                           & 1                                    & $43.65 (0.52)$ & $33.66 (0.58)$ & ---             \\
                                                                                               & 2                                    & $608.0 (6.8)$ & $717.3 (9.2)$ & ---             \\
                                                                                               & 3                                    & $7100 (160)$ & $13620 (310)$ & ---             \\ \hline
                \end{tabular}
                \egroup
            }
            \end{center}

@@

            \begin{center}
            \scalebox{0.75}{
                \bgroup
                \centering
                \def\arraystretch{1.5}
            \begin{tabular}{ccccc}
            \multirow{2}{*}{Feature}                                                       & \multirow{2}{*}{Dive / Subdive Type} & \multicolumn{3}{c}{Parameter Estimate}        \\
                                                                                           &                                      & $\hat \mu$    & $\hat \sigma$ & $\hat \phi$   \\ \hline
            \multirow{1}{*}{Dive Duration $(s)$ -- $Y_t$}                                  & 1                                    & $41.70 (1.24)$ & $30.50 (1.22)$ & ---           \\ \hline
            \multirow{3}{*}{$x$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_x$}   & 1                                    & $0.460 (0.030)$ & $0.039 (0.001)$ & $0.968 (0.002)$ \\
                                                                                           & 2                                    & $0.209 (0.012)$ & $0.096 (0.002)$ & $0.828 (0.007)$ \\
                                                                                           & 3                                    & $0.232 (0.035)$ & $0.296 (0.010)$ & $0.607 (0.023)$ \\ \hline
            \multirow{3}{*}{$y$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_y$}   & 1                                    & $0.457 (0.038)$ & $0.052 (0.001)$ & $0.968 (0.002)$ \\
                                                                                           & 2                                    & $0.436 (0.012)$ & $0.094 (0.002)$ & $0.828 (0.007)$ \\
                                                                                           & 3                                    & $0.365 (0.042)$ & $0.366 (0.012)$ & $0.607 (0.023)$ \\ \hline
            \multirow{3}{*}{$z$-Acc. $(m/s^2)$ -- $\left(\Zone_{t,\tilde t^*}\right)_z$}   & 1                                    & $-0.704 (0.044)$ & $0.059 (0.001)$ & $0.968 (0.002)$ \\
                                                                                           & 2                                    & $-0.571 (0.014)$ & $0.111 (0.002)$ & $0.828 (0.007)$ \\
                                                                                           & 3                                    & $-0.301 (0.042)$ & $0.355 (0.012)$ & $0.607 (0.023)$ \\ \hline
            \multirow{3}{*}{Wiggliness - $\Ztwo_{t,\tilde t^*}$}                           & 1                                    & $34.40 (0.37)$ & $23.33 (0.38)$ & ---           \\
                                                                                           & 2                                    & $495.0 (5.7)$ & $505.4 (6.9)$ & ---           \\
                                                                                           & 3                                    & $9190 (220)$ & $13590 (360)$ & ---           \\ \hline
            \end{tabular}
                \egroup
            }
            \end{center}

                \noindent Table \arabic{tablenum}: Estimates and predicted standard errors of emission parameters for dive duration ($Y_t$), acceleration ($\Zone_{t,\tilde t^*}$), and wiggliness ($\Ztwo_{t,\tilde t^*}$) of the killer whale case study data under the CarHMM-DFT. The parentheses refer to the estimated standard error of the parameter estimates derived from the inverse of the observed information matrix.
                \addtocounter{tablenum}{1}

@@

\end{document}
